---
title: "sudipta_bulk_seq"
output: html_document
date: "2025-12-03"
---
```{r}
library(limma)
library(edgeR)
library(tidyverse)
library(DESeq2)
```

#load the .tsv file
```{r}
count_matrix_sudipta <- read.delim("D:/sudipta_bulk_analysis/new_analysis/RSEM_gene_count_matrix_sudipta.tsv")
```
```{r}
count_matrix_sudipta <- separate(count_matrix_sudipta,col = "X", into = c("Ensemble_Id","Gene_Id"),sep ="_")

```

```{r}
count_matrix_sudipta
count_matrix_sudipta_new <- count_matrix_sudipta[, -c(1,1)]
colnames(count_matrix_sudipta_new)
head(count_matrix_sudipta_new)
column_matrix_no4 <- count_matrix_sudipta_new[ , 1:10]
column_matrix_no4

write.csv(column_matrix_no4, "D:/sudipta_bulk_analysis/new_analysis/new_sudipta_soumita/column_matrix_no4.csv", row.names = TRUE)
column_matrix_no4 <- read.csv("D:/sudipta_bulk_analysis/new_analysis/new_sudipta_soumita/column_matrix_no4.csv", header = TRUE)

```

```{r}

row.names(column_matrix_no4) <- column_matrix_no4$Gene_Id
```
```{r}
# ── Option 1: Make duplicates unique (RECOMMENDED) ───────────────────────
# Appends .1, .2 etc. to duplicate names — keeps all genes
rownames(column_matrix_no4) <- make.unique(column_matrix_no4$Gene_Id)
column_matrix_no4$Gene_Id <- NULL

# Check it worked
any(duplicated(rownames(column_matrix_no4)))  # Should return FALSE
```


```{r}
#sample_name <- c(
#  "GSD_1A_S160_L002.genes.results",
#  "GSD_1B_S151_L002.genes.results",
# "GSD_1C_S159_L002.genes.results",
#  "GSD_2A_S161_L002.genes.results",
#  "GSD_2B_S152_L002.genes.results",
#  "GSD_2C_S158_L002.genes.results",
#  "GSD_3A_S162_L002.genes.results",
#  "GSD_3B_S153_L002.genes.results",
#  "GSD_3C_S157_L002.genes.results",
#  "GSD_4A_S163_L002.genes.results",
#  "GSD_4B_S154_L002.genes.results",
#  "GSD_4C_S156_L002.genes.results"
#)
# Define conditions
#Condition <- c(
#  rep("Control", 4),   # first 4 samples are control
#  rep("DCZ_Treated", 8)    # remaining 8 samples are treated
#)
# Define treatment times
#Time <- c(
#  rep("0h", 4),  # controls
#  rep("1h", 4),  # DCZ 1 hour
#  rep("2h", 4)   # DCZ 2 hours
#)
samples <- colnames(column_matrix_no4)
Condition <- case_when(
  grepl("A", samples) ~ "Control",
  grepl("B", samples) ~ "DCZ_Treated_1H",
  grepl("C", samples) ~ "DCZ_Treated_2H"
)

Time <- ifelse(grepl("A", samples), "0h",
         ifelse(grepl("B", samples), "1h",
         ifelse(grepl("C", samples), "2h", NA)))

# Combine into a data frame
sudipta_sample <- data.frame(
  Sample_name = samples,
  Condition = Condition,
  Time = Time
)
# Set sample_name as rownames (required by DESeq2)
#sudipta_sample$sample_name <- NULL
sudipta_sample
# Save to CSV
write.csv(sudipta_sample, "D:/sudipta_bulk_analysis/new_analysis/sudipta_sample.csv", row.names = TRUE)
```


```{r}

column_matrix_no4 <- round(as.matrix(column_matrix_no4))
storage.mode(column_matrix_no4) <- "integer"
sudipta_sample_info <- read.csv("D:/sudipta_bulk_analysis/new_analysis/sudipta_sample.csv", header = TRUE)
head(sudipta_sample_info)
``` 

```{r}
sudipta_sample_info$Condition <- factor(sudipta_sample_info$Condition)
sudipta_sample_info$Time <- factor((sudipta_sample_info$Time))

```

#creating seseq object
```{r}
sudipta_sample_info$Condition <- factor(sudipta_sample_info$Condition)
dds <- DESeqDataSetFromMatrix(countData = column_matrix_no4, 
                               colData = sudipta_sample_info, design = ~ Condition)


```

```{r}
dds$Condition <- factor(dds$Condition, levels = c("Control","DCZ_Treated_1H", "DCZ_Treated_2H"))
```

## filter low quality gene
```{r}
keep <- rowSums(counts(dds)) >= 10
dds <-  dds[keep,]
```
#identify differentially expressed gene

```{r}
dds <- DESeq(dds)
#deseq result
deseq_results <- results(dds)
deseq_results
```

```{r}
deseq_results <- as.data.frame(deseq_results)
class(deseq_results)
deseq_results
```
#order the deseq_result on p-value
```{r}
deseq_results_ordered <- deseq_results[order(deseq_results$pvalue),]
deseq_results_ordered
```
## select gene with most differentially expressed
## filtered based on the p- adjusted value

```{r}
filtered_padj <- deseq_results %>% filter(deseq_results$padj < 0.05)
filtered_padj
```

#filter based on log2fc
```{r}
filtered_padj <- filtered_padj %>% filter(abs(filtered_padj$log2FoldChange) > 0.5)
filtered_padj
```

```{r}
dim(deseq_results)
dim(filtered_padj)
```


```{r}
write.csv(deseq_results, "D:/sudipta_bulk_analysis/new_analysis/new_sudipta_soumita/deseq_results_sudipta.csv", row.names = TRUE)
write.csv(filtered_padj, "D:/sudipta_bulk_analysis/new_analysis/new_sudipta_soumita/filtered_padj_1.csv", row.names = TRUE)
```

## normalized counts
```{r}
normalized_counts <- counts(dds, normalized = TRUE)
write.csv(normalized_counts, "D:/sudipta_bulk_analysis/new_analysis/new_sudipta_soumita/normalized_counts.csv", row.names = TRUE)
normalized_counts
```

### dispersion plot
```{r}
plotDispEsts(dds)
```
### pca plot(diamentionality reduction technique)
### variance stablizing transformation 
```{r}
vsd <- vst(dds,blind = FALSE)
```

```{r}
#plotPCA(vsd, intgroup=c("Control","DCZ_Treated_1H", "DCZ_Treated_2H"))
```


```{r}
# Make sure Condition is a factor with desired order
colData(vsd)$Condition <- factor(colData(vsd)$Condition,
                                 levels = c("Control","DCZ_Treated_1H","DCZ_Treated_2H"))

# Correct plotPCA call
plotPCA(vsd, intgroup = "Condition")

```


```{r}
# Ensure rownames match columns of counts
rownames(sudipta_sample_info) <- sudipta_sample_info$Sample_name

# Rebuild dds if needed
dds <- DESeqDataSetFromMatrix(countData = column_matrix_no4,
                              colData   = sudipta_sample_info,
                              design    = ~ Condition)

vsd <- vst(dds)  # or rlog(dds)

```


```{r}
library(ggplot2)

p <- plotPCA(vsd, intgroup = "Condition")  # color by Condition
p + geom_text(aes(label = Sample_name), vjust = -0.5, size = 6)

```


```{r}
library(DESeq2)
barplot(colSums(counts(dds)),
        names.arg = colnames(dds),
        las = 2, cex.names = 0.6,
        main = "Library sizes per sample")

```

## heat map of sample to sample distance matrix 
## genaetrate the distance matrix
```{r}
sampleDists <- dist(t(assay(vsd)))
sampleDiatMatrix <- as.matrix(sampleDists)
colnames(sampleDiatMatrix)
```

#SET a colour scheme
```{r}
library(RColorBrewer)   # for brewer.pal()
library(pheatmap)
colors <- colorRampPalette(rev(brewer.pal(100,"Blues")))(225)
```

## generat the heat map
```{r}
pheatmap(sampleDiatMatrix, clustering_distance_rows = sampleDists,
         clustering_distance_col = sampleDists, col = colors, scale = "row")
```
```{r}
png("D:/sudipta_bulk_analysis/new_analysis/new_sudipta_soumita/sample_heatmap.png",
    width = 2000, height = 2000, res = 300)

pheatmap(sampleDiatMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col = colors,
         scale = "row")

dev.off()
```



## heatmap of log tarnsformed normalizes
```{r}
top_hits <- deseq_results[order(deseq_results$padj), ][1:150,]
top_hits <- row.names(top_hits)
top_hits
```


```{r}
rld <- rlog(dds, blind = FALSE)


pheatmap(assay(rld)[top_hits,], cluster_row=T,show_rownames=TRUE,cluster_cols=T, cluster_rows=T, scale = "row")
#pheatmap(assay(rld)[top_hits,])
```


```{r}

annot_info <- as.data.frame(colData(dds)[,c('Condition','Time')])
pheatmap(assay(rld)[top_hits,],cluster_rows=TRUE,show_rownames=TRUE,cluster_cols=TRUE,
         annotation_col=annot_info, scale = "row")
         

```

```{r}
mat <- assay(rld)[top_hits, ]
colnames(mat) <- gsub("_S.*", "", colnames(mat))
colnames(mat)
```


```{r}
mat_scaled <- t(scale(t(mat)))

# clip extremes
mat_scaled[mat_scaled >  3] <-  3
mat_scaled[mat_scaled < -3] <- -3
```

```{r}
desired_order <- c(
  "GSD_3A","GSD_2A","GSD_1A",
  "GSD_3B","GSD_2B","GSD_1B",
  "GSD_3C","GSD_2C","GSD_1C"
)
setdiff(desired_order, colnames(mat_scaled))
```

```{r}
mat_scaled <- mat_scaled[, desired_order]
```

```{r}
pheatmap(mat_scaled,
         cluster_rows = TRUE,
         cluster_cols = FALSE,   #keeps your order
         show_rownames = TRUE,
         scale = "none")
```


```{r}
pheatmap(mat_scaled,
         cluster_rows = TRUE,
         cluster_cols = FALSE,
         gaps_col = c(3,6),   # separates A | B | C
         show_rownames = TRUE,
         scale = "none")
```
```{r}
# ── Annotation
annot_info <- as.data.frame(colData(dds)[, c("Condition","Time")])

# IMPORTANT: match annotation order with matrix
annot_info <- annot_info[colnames(mat_scaled), ]

# ── Heatmap 
pheatmap(mat_scaled,
         # clustering
         cluster_rows = TRUE,
         cluster_cols = FALSE,   # keep your custom order

         # gaps for A | B | C grouping
         gaps_col = c(3, 6),

         # annotation
         annotation_col = annot_info,

         # labels
         show_rownames = TRUE,
         show_colnames = TRUE,

         # scaling (already scaled)
         scale = "none")
```


```{r}
png(filename = "D:/sudipta_bulk_analysis/new_analysis/new_sudipta_soumita/heat_map_1.png",res = 400,units = "in",height = 15,width = 10)
pheatmap(mat_scaled,
         # clustering
         cluster_rows = TRUE,
         cluster_cols = FALSE,   # keep your custom order

         # gaps for A | B | C grouping
         gaps_col = c(3, 6),

         # annotation
         annotation_col = annot_info,

         # labels
         show_rownames = TRUE,
         show_colnames = TRUE,

         # scaling (already scaled)
         scale = "none")
dev.off()
```


```{r}
```


```{r}
# reorder matrix
mat_scaled <- mat_scaled[, desired_order]

# reorder annotation
annot_info <- annot_info[desired_order, ]

# reorder labels
col_labels <- col_labels[match(desired_order, colnames(mat_scaled))]
```


```{r}
```


```{r}
```


```{r}
```

# heatmap of the z-score
```{r}
cal_z_score <- function(x) { (x - mean(x)) / sd(x) }

Zscore_all <- t(apply(normalized_counts, 1, cal_z_score))

zscore_subset <- Zscore_all[top_hits, ]

pheatmap(zscore_subset)
```

##ma plot 
```{r}
plotMA(dds, ylim = c(-2, 2))

dds <- DESeq(dds)
res <- results(dds, alpha = 0.05)  # FDR threshold = 0.05
plotMA(res, ylim = c(-2, 2))

```


```{r}
#reset par
par(mfrow=c(1,1))
# Make a basic volcano plot(padj)
with(filtered_padj, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-3,3)))

# Add colored points: blue if padj<0.01, red if log2FC>1 and padj<0.05)
with(subset(filtered_padj, padj<.01 ), points(log2FoldChange, -log10(pvalue), pch=20, col="blue"))
with(subset(filtered_padj, padj<.01 & abs(log2FoldChange)>2), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```


```{r}
#reset par
#par(mfrow=c(1,1))
#with(normalized_counts, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-3,3)))

# Add colored points: blue if padj<0.01, red if log2FC>1 and padj<0.05)
#with(subset(normalized_counts, padj<.01 ), points(log2FoldChange, -log10(pvalue), pch=20, col="blue"))
#with(subset(normalized_counts, padj<.01 & abs(log2FoldChange)>2), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```


```{r}
# Run DESeq if not already done
dds <- DESeq(dds)

# Get results with adjusted p-values
res <- results(dds, alpha = 0.01)

# Order by padj (optional, for clarity
res_ordered <- res[order(res$padj), ]
res_ordered
# Basic volcano plot
with(res_ordered,
     plot(log2FoldChange, -log10(pvalue),
          pch = 20,
          main = "Volcano Plot",
          xlim = c(-10, 15),
          xlab = "log2 Fold Change",
          ylab = "-log10(p-value)",
          col = "gray50"))

# Highlight significant genes
with(subset(res_ordered, padj < 0.01 & abs(log2FoldChange) > 1),
     points(log2FoldChange, -log10(pvalue), pch = 20, col = "red"))

# Add gene names for top hits
topgenes <- subset(res_ordered, padj < 0.01 & abs(log2FoldChange) > 1)
topgenes <- topgenes[order(topgenes$padj), ][1:50, ]

text(topgenes$log2FoldChange, -log10(topgenes$pvalue),
     labels = rownames(topgenes), cex = 0.7, pos = 3, col = "black")

```

### remove noiice
```{r}
resultsNames(dds)
```


```{r}
library(apeglm)
```

## 1-hour DCZ-treated vs Control
```{r}
reslfc_1h <- lfcShrink(
  dds,
  coef = "Condition_DCZ_Treated_1H_vs_Control",
  type = "apeglm"
)

plotMA(reslfc_1h, ylim=c(-2,2), main="DCZ 1h vs Control")
```
```{r}
library(DESeq2)
library(ggplot2)
library(ggrepel)  
```

```{r}
# Convert to data frame and clean
resdf_1 <- as.data.frame(reslfc_1h)
resdf_1 <- na.omit(resdf_1
                   )
write.csv(resdf_1, "D:/sudipta_bulk_analysis/new_analysis/resdf_1_1h_vs_C.csv", row.names = TRUE)
# Add significance column
resdf_1$significance <- "Not Sig"
resdf_1$significance[resdf_1$padj < 0.05 & abs(resdf_1$log2FoldChange) > 1] <- "Significant"

# Select top genes for labeling
top_genes <- rownames(resdf_1[resdf_1$padj < 0.01 & abs(resdf_1$log2FoldChange) > 2, ])

# Volcano plot
ggplot(resdf_1, aes(x = log2FoldChange, y = -log10(padj))) +
  geom_point(aes(color = significance), alpha = 0.7, size = 2) +
  scale_color_manual(values = c("grey", "red")) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "blue") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "blue") +
  geom_text_repel(aes(label = ifelse(rownames(resdf_1) %in% top_genes, rownames(resdf_1), "")),
                  size = 3, max.overlaps = 10) +
  labs(title = "Volcano Plot: Control vs DCZ Treated (1h)",
       x = "log2 Fold Change",
       y = "-log10(adjusted p-value)",
       color = "Significance") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top")

```


## 2-hour DCZ-treated vs Control
```{r}
reslfc_2h <- lfcShrink(
  dds,
  coef = "Condition_DCZ_Treated_2H_vs_Control",
  type = "apeglm"
)
plotMA(reslfc_2h, ylim=c(-2,2), main="DCZ 2h vs Control")
```


```{r}
# Convert to data frame and clean
resdf_2 <- as.data.frame(reslfc_2h)
resdf_2 <- na.omit(resdf_2)
write.csv(resdf_2, "D:/sudipta_bulk_analysis/new_analysis/resdf_2_2h_vs_C.csv", row.names = TRUE)

view(resdf_2)
resdf_2$hehe <- rownames(resdf_2)
resdf_2 <- separate(resdf_2,col = "hehe",into = c("ensemble_id","gene_name"),sep = "_")

# Add significance column
resdf_2$significance <- "Not Sig"
resdf_2$significance[resdf_2$padj < 0.05 & abs(resdf_2$log2FoldChange) > 1] <- "Significant"

# Select top genes for labeling
top_genes <- rownames(resdf_2[resdf_2$padj < 0.01 & abs(resdf_2$log2FoldChange) > 2, ])

# Volcano plot for 2h
ggplot(resdf_2, aes(x = log2FoldChange, y = -log10(padj))) +
  geom_point(aes(color = significance), alpha = 0.7, size = 2) +
  scale_color_manual(values = c("grey", "red")) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "blue") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "blue") +
  geom_text_repel(aes(label = ifelse(rownames(resdf_2) %in% top_genes, rownames(resdf_2), "")),
                  size = 3, max.overlaps = 10) +
  labs(title = "Volcano Plot: Control vs DCZ Treated (2h)",
       x = "log2 Fold Change",
       y = "-log10(adjusted p-value)",
       color = "Significance") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top")


```


```{r}
library(EnhancedVolcano)

png(filename = "D:/sudipta_bulk_analysis/new_analysis/Contro vs DCZ 2h.png",res = 400,units = "in",height = 10,width = 15)
# 1️⃣  DCZ 1h vs Control
EnhancedVolcano(resdf_2,
    lab = resdf_2$gene_name,
    x = 'log2FoldChange',
    y = 'pvalue',
    title = 'Control vs DCZ Treated (2h)',
    subtitle = 'Bulk RNA-seq Differential Expression',
    caption = 'Threshold: log2FC > 0.5 | p < 0.05',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    pointSize = 2.0,
    labSize = 3.0,
    col = c('grey30', 'royalblue', 'forestgreen', 'red2'),
    colAlpha = 0.8,
    legendPosition = 'top',
    drawConnectors = TRUE,
    widthConnectors = 0.3,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    border = 'full'
)
dev.off()

```
```{r}
resdf_1$hehe <- rownames(resdf_1)
resdf_1 <- separate(resdf_1,col = "hehe",into = c("ensemble_id","gene_name"),sep = "_")
view(res_df_1)
```

```{r}
png(filename = "D:/sudipta_bulk_analysis/new_analysis/Contro vs DCZ 1h.png",res = 400,units = "in",height = 10,width = 15)
# 1️⃣  DCZ 1h vs Control
EnhancedVolcano(resdf_1,
    lab = resdf_1$gene_name,
    x = 'log2FoldChange',
    y = 'pvalue',
    title = 'Control vs DCZ Treated (1h)',
    subtitle = 'Bulk RNA-seq Differential Expression',
    caption = 'Threshold: log2FC > 1 | p < 0.05',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    pointSize = 2.0,
    labSize = 3.0,
    col = c('grey30', 'royalblue', 'forestgreen', 'red2'),
    colAlpha = 0.8,
    legendPosition = 'top',
    drawConnectors = TRUE,
    widthConnectors = 0.3,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    border = 'full'
)
dev.off()
```
```{r}

```


```{r}
BiocManager::install("clusterProfiler")
BiocManager::install("pathview")
BiocManager::install("enrichplot")
library(clusterProfiler)
library(enrichplot)


organism = "org.Mm.eg.db"
BiocManager::install(organism, character.only = TRUE)
library(organism, character.only = TRUE)

keytypes(org.Mm.eg.db)
```


```{r}
resdf_2_sig <- resdf_2[resdf_2$padj < 0.05 & resdf_2$log2FoldChange > 1 | resdf_2$log2FoldChange < -1,]
```


```{r}
#stat_genes <- resdf_2_sig$log2FoldChange
#names(stat_genes) <- resdf_2_sig$gene_name
#stat_genes <- sort(stat_genes,decreasing = T)
#stat_genes_clean <- stat_genes[!duplicated(names(stat_genes))]

```


```{r}
install.packages("enrichR")
library("enrichR")


```
## pathway analysis betn control and dcz_2h

```{r}

resdf_2$Gene_symbol <- substr(resdf_2$Gene_symbol, 1, 18)
```


```{r}
view(resdf_2)

resdf_2$hehe <- rownames(resdf_2)
resdf_2 <- separate(resdf_2,col = "hehe",into = c("Gene_Id","Gene_symbol"),sep = "_")
```



# between control vs DCZ_2h #upregulated #downregulated
```{r}
up.idx <- which(resdf_2$padj < 0.05 & resdf_2$log2FoldChange > 0.5) # FDR < 0.05 and logFC > 0
dn.idx <- which(resdf_2$padj < 0.05 & resdf_2$log2FoldChange < -0.5) # FDR < 0.05 and logFC < 0

```


```{r}
dim(resdf_2)
length(up.idx)
length(dn.idx)
```


```{r}
up.genes <- resdf_2[up.idx,]$Gene_symbol
dn.genes <- resdf_2[dn.idx,]$Gene_symbol
```


```{r}
head(up.genes, 100)
```


```{r}
head(dn.genes, 100)
```
```{r}
dbs <- listEnrichrDbs()
dbs <- dbs[order(dbs$libraryName),]

class(dbs)
```
```{r}
dim(dbs)
head(dbs)
dbs$libraryName
dbs
```

## Select databases
```{r}
dbs_go <- c("GO_Molecular_Function_2018", "GO_Cellular_Component_2018", "GO_Biological_Process_2018")
dbs_pw <- c("KEGG_2019_Mouse", "WikiPathways_2019_Mouse", "BioPlanet_2019")

```

```{r}
library(clusterProfiler)
library(org.Mm.eg.db)

keytypes(org.Mm.eg.db)

```

### Perform enrichment analysis
## GO analysis
```{r}
## upregulatiing
upEnriched_go <- enrichr(genes = up.genes, databases = dbs_go)
```


```{r}
## downregulating
dnEnriched_go <- enrichr(genes = dn.genes, databases = dbs_go)
```


```{r}
upEnriched_pw <- enrichr(genes = up.genes, databases = dbs_pw)
```


```{r}
dnEnriched_pw <- enrichr(genes = dn.genes, databases = dbs_pw)
```


```{r}
png(filename = "D:/sudipta_bulk_analysis/new_analysis/bioplanet_up_pathways(.png",res = 400,units = "in",width = 8,height = 10)
plotEnrich(upEnriched_pw[[3]], showTerms = 25, numChar = 80, y = "Count", orderBy = "P.value")
dev.off()
```
```{r}
png(filename = "D:/sudipta_bulk_analysis/new_analysis/kegg_dn_pathways.png",res = 400,units = "in",width = 8,height = 8)
plotEnrich(dnEnriched_pw[[1]], showTerms = 25, numChar = 80, y = "Count", orderBy = "P.value")
dev.off()
```
```{r}
png(filename = "D:/sudipta_bulk_analysis/new_analysis/KEGG_pathways_upregulated_2h.png",
    res = 400, units = "in", width = 8, height = 10)

plotEnrich(upEnriched_pw[[1]],
           showTerms = 25,
           numChar = 80,
           y = "Count",
           orderBy = "P.value") +
  ggtitle("Control vs DCZ_treated_2H_Upregulated") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )

dev.off()

```
downregulated
```{r}
png(filename = "D:/sudipta_bulk_analysis/new_analysis/WikiPathways_pathways_upregulated_2h.png",
    res = 400, units = "in", width = 8, height = 10)

plotEnrich(dnEnriched_pw[[2]],
           showTerms = 25,
           numChar = 80,
           y = "Count",
           orderBy = "P.value") +
  ggtitle("Control vs DCZ_treated_2H_Upregulated") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )

dev.off()
```


```{r}
#write.csv(upEnriched_pw[[1]], file = "D:/sudipta_bulk_analysis/new_analysis/KEGG_pathways_up_2h.csv", row.names = FALSE)
```
```{r}
# Write each enrichment table to a separate CSV file
write.csv(upEnriched_pw[[1]], 
          file = "D:/sudipta_bulk_analysis/new_analysis/KEGG_pathway_upregulated_2h.csv",
          row.names = FALSE)

write.csv(upEnriched_pw[[2]], 
          file = "D:/sudipta_bulk_analysis/new_analysis/WikiPathways_pathways_upregulated_2h.csv",
          row.names = FALSE)

write.csv(upEnriched_pw[[3]], 
          file = "D:/sudipta_bulk_analysis/new_analysis/BioPlanet_up_pathway_upregulated_2h.csv",
          row.names = FALSE)
```

## pathway analysis betn control and dcz_1h 
```{r}
view(resdf_1)

resdf_1$hehe <- rownames(resdf_1)
resdf_1 <- separate(resdf_1,col = "hehe",into = c("Gene_Id","Gene_symbol"),sep = "_")

resdf_1$Gene_symbol <- substr(resdf_1$Gene_symbol, 1, 18)

```



#upregulated #downregulated
```{r}
up.idx <- which(resdf_1$padj < 0.05 & resdf_1$log2FoldChange > 0.5) # FDR < 0.05 and logFC > 0
dn.idx <- which(resdf_1$padj < 0.05 & resdf_1$log2FoldChange < -0.5) # FDR < 0.05 and logFC < 0

```


```{r}
dim(resdf_1)
length(up.idx)
length(dn.idx)
```


```{r}
up.genes <- resdf_1[up.idx,]$Gene_symbol
dn.genes <- resdf_1[dn.idx,]$Gene_symbol
```


```{r}
## upregulatiing
upEnriched_go <- enrichr(genes = up.genes, databases = dbs_go)
## downregulatiing
dnEnriched_go <- enrichr(genes = dn.genes, databases = dbs_go)
```


```{r}
#upregulating
upEnriched_pw <- enrichr(genes = up.genes, databases = dbs_pw)
#doenregulating
dnEnriched_pw <- enrichr(genes = dn.genes, databases = dbs_pw)
```


```{r}
png(filename = "D:/sudipta_bulk_analysis/new_analysis/BioPlanet_up_pathway_upregulated_1h.png",
    res = 400, units = "in", width = 8, height = 10)

plotEnrich(upEnriched_pw[[3]],
           showTerms = 25,
           numChar = 80,
           y = "Count",
           orderBy = "P.value") +
  ggtitle("Control vs DCZ_treated_1H_Upregulated") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )

dev.off()
write.csv(upEnriched_pw[[3]], 
          file = "D:/sudipta_bulk_analysis/new_analysis/BioPlanet_up_pathway_upregulated_1h.csv",
          row.names = FALSE)
```


```{r}

```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
library(fgsea)
#imporing all gmt pathway files using fgsea package downloaded from Molecular signature database
cannonical_pathways <- gmtPathways("D:/sudipta_bulk_analysis/gmt files/cannonical.Mm.symbols.gmt")

hallmark_gene_sets <- gmtPathways("D:/sudipta_bulk_analysis/gmt files/hallmark.symbols.gmt")

ontology_gene_sets <- gmtPathways("D:/sudipta_bulk_analysis/gmt files/gene ontology..Mm.symbols.gmt")
```

### ontology_gene_set
```{r}
sig_genes <- resdf_KO %>% 
    filter(padj < 0.05) %>% 
    filter(log2FoldChange > 0.25 | log2FoldChange < -0.25)

sig_genes <- sig_genes %>% dplyr::arrange(desc(log2FoldChange))

stat_genes <- sig_genes$log2FoldChange
names(stat_genes) <- sig_genes$Gene_symbol

stat_genes <- stat_genes[!duplicated(names(stat_genes))]

```



```{r}
ontology_gene_set_pathway <- fgsea(
                              pathways = ontology_gene_sets,
                              stats = stat_genes,
                              minSize = 3,
                              maxSize = 800,
                              nproc = 1)


convert_list_columns <- function(df) {
 for (col in names(df)) {
    if (is.list(df[[col]])) {
      df[[col]] <- sapply(df[[col]], toString)
    }
  }
  return(df)
}
ontology_gene_set_pathway <- convert_list_columns(ontology_gene_set_pathway)
write.csv(ontology_gene_set_pathway,"D:/sudipta_bulk_analysis/new_analysis/fgsea_ontology_gene_set_pathway_2H.csv")
```


### cannonical_pathways
```{r}
cannonical_gene_set_pathway <- fgsea(
                              pathways = cannonical_pathways,
                              stats = stat_genes,
                              minSize = 3,
                              maxSize = 800,
                              nproc = 1)


convert_list_columns <- function(df) {
 for (col in names(df)) {
    if (is.list(df[[col]])) {
      df[[col]] <- sapply(df[[col]], toString)
    }
  }
  return(df)
}
cannonical_gene_set_pathway <- convert_list_columns(cannonical_gene_set_pathway)
write.csv(cannonical_gene_set_pathway,"D:/sudipta_bulk_analysis/new_analysis/fgsea_cannonical_gene_set_pathway_2H.csv")
```

##hallmark_gene_sets
```{r}
hallmark_gene_sets_pathway <- fgsea(
                              pathways = hallmark_gene_sets,
                              stats = stat_genes,
                              minSize = 3,
                              maxSize = 800,
                              nproc = 1)


convert_list_columns <- function(df) {
 for (col in names(df)) {
    if (is.list(df[[col]])) {
      df[[col]] <- sapply(df[[col]], toString)
    }
  }
  return(df)
}
hallmark_gene_sets_pathway <- convert_list_columns(hallmark_gene_sets_pathway)
write.csv(hallmark_gene_sets_pathway,"D:/sudipta_bulk_analysis/new_analysis/fgsea_hallmark_gene_sets_pathway_2H.csv")
```




for control vs 1 hrs
```{r}
sig_genes <- resdf_1 %>% 
    filter(padj < 0.05) %>% 
    filter(log2FoldChange > 0.5 | log2FoldChange < -0.5)

sig_genes <- sig_genes %>% dplyr::arrange(desc(log2FoldChange))

stat_genes <- sig_genes$log2FoldChange
names(stat_genes) <- sig_genes$Gene_symbol

stat_genes <- stat_genes[!duplicated(names(stat_genes))]
```

#ontology_pathway
```{r}
ontology_gene_set_pathway <- fgsea(
                              pathways = ontology_gene_sets,
                              stats = stat_genes,
                              minSize = 3,
                              maxSize = 800,
                              nproc = 1)


convert_list_columns <- function(df) {
 for (col in names(df)) {
    if (is.list(df[[col]])) {
      df[[col]] <- sapply(df[[col]], toString)
    }
  }
  return(df)
}
ontology_gene_set_pathway <- convert_list_columns(ontology_gene_set_pathway)
write.csv(ontology_gene_set_pathway,"D:/sudipta_bulk_analysis/new_analysis/fgsea_ontology_gene_set_pathway_1H.csv")
```
## pathway cannonical

```{r}
cannonical_gene_set_pathway <- fgsea(
                              pathways = cannonical_pathways,
                              stats = stat_genes,
                              minSize = 3,
                              maxSize = 800,
                              nproc = 1)


convert_list_columns <- function(df) {
 for (col in names(df)) {
    if (is.list(df[[col]])) {
      df[[col]] <- sapply(df[[col]], toString)
    }
  }
  return(df)
}
cannonical_gene_set_pathway <- convert_list_columns(cannonical_gene_set_pathway)
write.csv(cannonical_gene_set_pathway,"D:/sudipta_bulk_analysis/new_analysis/fgsea_cannonical_gene_set_pathway_1H.csv")
```
## hallmark_gene_sets_pathway

```{r}
hallmark_gene_sets_pathway <- fgsea(
                              pathways = hallmark_gene_sets,
                              stats = stat_genes,
                              minSize = 3,
                              maxSize = 800,
                              nproc = 1)


convert_list_columns <- function(df) {
 for (col in names(df)) {
    if (is.list(df[[col]])) {
      df[[col]] <- sapply(df[[col]], toString)
    }
  }
  return(df)
}
hallmark_gene_sets_pathway <- convert_list_columns(hallmark_gene_sets_pathway)
write.csv(hallmark_gene_sets_pathway,"D:/sudipta_bulk_analysis/new_analysis/fgsea_hallmark_gene_sets_pathway_1H.csv")
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


